---
title: "Introduction to Text Analysis in R (Activity 2)"
author: "Christian Staal Bruun Overgaard"
date: "April 25, 2023"
output: html_document
---

*This file doesn't have the solutions included. The solutions are provided in a seperate file.*

In this tutorial, you'll get the chance to use the skills you learned about in *Activity 1*. Please feel free to go through the exercises below with a dataset of your own.

## Load packages

```{r load_packages}
library(readr) #for reading csv's
library(readxl) #for reading Excel sheets 
library(tidytext) #main package for text analysis
library(tidyverse) #collection of useful packages
library(ggthemes) #for customizing plots
library(wordcloud) 
```

## Import Data

In this tutorial, we'll work with a dataset that consists of 30,000 tweets about COVID-19, posted between January and May in 2020. (Or, you can work with a dataset of your own if you'd like!)

### Load

First, load either (A) Your own data, or (B) the file, "10K_Facebook_posts_for_tutorial_USnewsoutlets.csv" into R:

```{r }

```

### Inspect

Do a quick inspection of the data. TIP: You can use the `head()` function or the `glimpse()` function.

```{r}

```

### Select the relevant columns

Some of the variables consistent of two or more words seperated by blank spaces. That's quite inconvenient; in R, life is easier when there are no blank spaces in our variable names. Fortunately, the `janitor` package makes it easy to standardize the column names in a given dataset. (I encourage you to check out the `janitor` package but feel free to proceed without it if you prefer.)

```{r}

```

In this tutorial, we'll only use the id and Message colums:

```{r}

```

## Subsetting data

### Use words to subset

Try to subset to include only messages that mention the word "President"

```{r}

```

Now, try to subset to messages that include either the word "President" OR the word "Biden":

```{r}

```

### Bonus: Looking at only specific Facebook pages

We didn't talk about this in the first tutorial but if you wanted to only view content from a specific Facebook page, this is how you could do it:

```{r}
df_thehill <- df %>%
  filter(user_name == "TheHill")
```

* This kinds of filtering can be useful. The plots we're doing below could, for example, be done for different news outlets. 

In this tutorial, we'll analyze the full dataset (so we will use the dataframe called "df").

## Working with the TidyText package

### Transform data to relevant structures

#### Tokenizing

First, use the function, `unnest_tokens` to tokenize the data. Remember, the function takes two arguments: level you want to tokenize at (use "word") and the name of the column with the text (in our case "message").

```{r}

```

Now, do a count to identify the most common words. (HINT: It's totally fine to look at the code from Activity 1 or from any other source.)

```{r}

```

#### Removing "stopwords"

It's hard to gauge the meaning of the content, with all these stop words, right? Let's try to remove them! (HINT: you can use the anti_join function.)

```{r}

```

Remove any additional words that you think would be helpful to remove:

```{r}

```

Now, do a final count of the top words:

```{r}

```

### Visualizing common words

#### Bar plot

Next, make the initial preperation for the ggplot (HINT: use the functions, count, filter, mutate—encouraged to look at the code from Activity 1).

How many times do you think a word needs to appear to be worth including in the bar plot? (HINT: use the table above as a reference point when making that decision.)

```{r}


```

Then,  build the basic ggplot:

```{r}


```

Try making the plot look nicer by adding some labels and a theme:

```{r}


```


##### Wordcloud

As the final step, try making a word cloud:

```{r}


```


